### 2025-06-23 manuscript/main.tex
finsihed bnf inclusion

\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{bibliography/theory.bib}
\addbibresource{bibliography/implementation.bib}

\def\code#1{\texttt{#1}}

@@

Each approach has its benefits and trade-offs, and the choice depends upon the specific requirements, constraints, and long-term maintainability of the language tooling in question.

\chapter{Langium}
\section{High-Level Architecture Diagram}
\section{Module Decomposition}
\section{Data Flow and Control Flow}
\section{Language Server Protocol}

@@

\chapter{Testing \& Validation}
\section{Unit Tests}
\section{Integration Tests}
\section{Case Studies \& Examples}

%=== Part III - Evaluation \& Discussion ===%
\part{Evaluation \& Discussion}

@@

\chapter{Discussion?}
\section{Meeting the Requirements}

@@

\begin{figure}[htbp]
  \centering
  \begin{minipage}[t]{0.48\textwidth}
    \[
      \begin{array}{rcl}
        prog
         & \to  & (decl\;\texttt{;})^*\;\;func\;func\_list?                                            \\[6pt]

        decl\_list
         & \to  & decl\;\texttt{;}\;decl\_list
        \;\mid\; \epsilon                                                                              \\[6pt]

        func\_list
         & \to  & func\;func\_list
        \;\mid\; \epsilon                                                                              \\[6pt]

        decl
         & \to  & type\;\mathit{id}\;id\_list                                                          \\[6pt]

        id\_list
         & \to  & \texttt{,}\;\mathit{id}\;id\_list
        \;\mid\; \epsilon                                                                              \\[6pt]

        func
         & \to  & \mathit{id}\;\texttt{(}
        [\,type\;\texttt{\&}?\;\mathit{id}
          (\texttt{,}\;type\;\texttt{\&}?\;\mathit{id})^*\,]?
        \texttt{)}                                                                                     \\
         &      & \texttt{\{}(decl\;\texttt{;})^*\;block\;\texttt{\}}                                  \\[6pt]

        block
         & \to  & stmt\;stmt\_list                                                                     \\[6pt]

        block\_opt
         & \to  & block
        \;\mid\; \epsilon                                                                              \\[6pt]

        stmt\_list
         & \to  & stmt\;stmt\_list
        \;\mid\; \epsilon                                                                              \\[6pt]

        stmt
         & \to  & lval\;\texttt{=}\;expr\;\texttt{;}
        \;\mid\; lval\;\texttt{=}\;\texttt{Distribution(...)}\;\texttt{;}                              \\
         & \mid & \texttt{query}\;func\_call\;\texttt{;}
        \;\mid\; \texttt{observe}\;expr\;\texttt{;}                                                    \\
         & \mid & func\_call\;\texttt{;}
        \;\mid\; \texttt{throw}\;\texttt{;}                                                            \\
         & \mid & \texttt{if}\;\texttt{(}expr\texttt{)}\;\{\;stmt^*\;\}\;\texttt{else}\;\{\;stmt^*\;\} \\
         & \mid & \texttt{while}\;\texttt{(}expr\texttt{)}\;\{\;stmt^*\;\}                             \\
         & \mid & \texttt{try}\;\{\;stmt^*\;\}\;\texttt{catch}\;\{\;stmt^*\;\}                         \\[6pt]

        assignment
         & \to  & lval\;\texttt{=}\;assign\_rhs                                                        \\[6pt]

        assign\_rhs
         & \to  & expr
        \;\mid\; distribution                                                                          \\[6pt]

        func\_call
         & \to  & \mathit{id}\;\texttt{(}
        (expr\;(\texttt{,}\;expr)^*)?
        \texttt{)}                                                                                     \\[6pt]

        arg\_list\_opt
         & \to  & arg\_list
        \;\mid\; \epsilon                                                                              \\[6pt]

        arg\_list
         & \to  & arg\;arg\_list\_tail                                                                 \\[6pt]

        arg\_list\_tail
         & \to  & \texttt{,}\;arg\;arg\_list\_tail
        \;\mid\; \epsilon                                                                              \\[6pt]

        arg
         & \to  & expr                                                                                 \\[6pt]

        lval
         & \to  & \mathit{id}
        \;\mid\; \mathit{id}\;\texttt{[}expr\texttt{]}                                                 \\[6pt]

        distribution
         & \to  & \texttt{Bernoulli(}expr\texttt{,}expr\texttt{)}
        \;\mid\; \texttt{Uniform(}expr\texttt{,}expr\texttt{)}                                         \\[6pt]

        param
         & \to  & type\;\texttt{\&}?\;\mathit{id}                                                      \\[6pt]

        param\_list
         & \to  & param\;param\_list\_tail                                                             \\[6pt]

        param\_list\_tail
         & \to  & \texttt{,}\;param\;param\_list\_tail
        \;\mid\; \epsilon                                                                              \\[6pt]

        expr
         & \to  & logical\_or\;prob\_assign\_opt                                                       \\[6pt]

        prob\_assign\_opt
         & \to  & prob\_assign
        \;\mid\; \epsilon                                                                              \\[6pt]
        completion
         & \to  & factor\;(\texttt{+}\;\mid\;\texttt{-})\;factor^*                                     \\[6pt]

        factor
         & \to  & division\;(\texttt{*}\;division)^*                                                   \\[6pt]

        division
         & \to  & unary\;((\texttt{/}\;\mid\;\texttt{\%})\;unary)^*                                    \\[6pt]

        unary
         & \to  & (\texttt{!}\;\mid\;\texttt{-})\;unary
        \;\mid\; primary                                                                               \\[6pt]

        primary
         & \to  & bool
        \;\mid\; int\_literal
        \;\mid\; lval
        \;\mid\; \texttt{(}expr\texttt{)}                                                              \\[6pt]

        identifier
         & \to  & letter\;ident\_tail                                                                  \\[6pt]

        ident\_tail
         & \to  & letter\_or\_digit\;ident\_tail
        \;\mid\; \epsilon                                                                              \\[6pt]

        letter
         & \to  & \texttt{a} \;\mid\; \dots \;\mid\; \texttt{Z} \;\mid\; \texttt{\_}                   \\[6pt]

        letter\_or\_digit
         & \to  & letter \;\mid\; digit                                                                \\[6pt]

        digit
         & \to  & \texttt{0} \;\mid\; \dots \;\mid\; \texttt{9}                                        \\[6pt]

        type
         & \to  & \texttt{bool}
        \;\mid\; int\_spec
        \;\mid\; int\_spec\;\texttt{[}int\_literal\texttt{]}                                           \\[6pt]

        digit\_seq
         & \to  & digit\;digit\_seq
        \;\mid\; digit                                                                                 \\[6pt]

        int\_spec
         & \to  & \texttt{u}\;digit\_seq
        \;\mid\; \texttt{s}\;digit\_seq                                                                \\[6pt]

        int\_literal
         & \to  & digit\_seq\;int\_spec                                                                \\[6pt]

        bool
         & \to  & \texttt{true} \;\mid\; \texttt{false}
      \end{array}
    \]
  \end{minipage}
  \caption{Full grammar of MiniProb DSL}
  \label{fig:bnf_grammar}
\end{figure}

>>>>

\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{bibliography/theory.bib}
\addbibresource{bibliography/implementation.bib}
\usepackage{./simplebnf}
\def\code#1{\texttt{#1}}

@@

\section{Data Flow and Control Flow}
\section{Language Server Protocol}

@@

@@

@@

\begin{bnf}
  $prog$ : \textsf{Program} ::=
  | $decl\_list\; func\; func\_list$
  ;;
  $decl$ : \textsf{Declaration} ::=
  | $type\; identifier\; id\_list$
  | $\epsilon$
  ;;
  $decl\_list$ ::=
  | $decl\; \code{;}\; decl\_list$
  | $\epsilon$
  ;;
  $id\_list$ ::=
  | \code{,}\; $identifier\; id\_list$
  | $\epsilon$
  ;;
  $func$ : \textsf{Function} ::=
  | $identifier\; \left(\; param\_list\; \right)\; \left\{\; decl\_list\; block\; \right\}$
  ;;
  $func\_call$ : \textsf{Func\-Call} ::=
  | $identifier\; \left(\; arg\_list\; \right)$
  ;;
  $arg$ : \textsf{Argument} ::=
  | $expr$
  ;;
  $arg\_list$ ::=
  | $arg\; arg\_list\_tail$
  | $\epsilon$
  ;;
  $arg\_list\_tail$ ::=
  | \code{,} $arg\; arg\_list\_tail$
  | $\epsilon$
  ;;
  $block$ : \textsf{Codeblock} ::=
  | $stmt\; stmt\_list$
  ;;
  $block\_opt$ ::=
  | $block$
  | $\epsilon$
  ;;
  $stmt$ : \textsf{Statement} ::=
  | $assignment\;$ \code{;}
  | $query\;$ \code{;}
  | $observation\;$ \code{;}
  | $func\_call\;$ \code{;}
  | \code{throw} \code{;}
  | $if\_then\_else$
  | $while$
  | $try\_catch$
  ;;
  $stmt\_list$ ::=
  | $stmt\; stmt\_list$
  | $\epsilon$
  ;;
  $assignment$ ::=
  | $lval\; \code{=}\; assign\_rhs$
  ;;
  $assign\_rhs$ ::=
  | $expr$
  | $distribution$
  ;;
  $query$ : \textsf{Query} ::=
  | \code{query} $func\_call$
  ;;
  $observation$ : \textsf{Conditioning} ::=
  | \code{observe} $expr$
  ;;
  $if\_then\_else$ : \textsf{IfThenElse} ::=
  | \code{if} $\left(\; expr\; \right) \left\{\; block\_opt\; \right\}$ \code{else} $\left\{\; block\_opt\; \right\}$
  ;;
  $while$ : \textsf{While-loop} ::=
  | \code{while} $\left(\; expr\; \right) \left\{\; block\_opt\; \right\}$
  ;;
  $try\_catch$ : \textsf{Try-catch}::=
  | \code{try} $\left\{\; block\_opt\; \right\} \code{catch} \left\{\; block\_opt\; \right\}$
  ;;
  $lval$ ::=
  | $identifier\;\left[\;expr\;\right]$
  | $identifier$
  ;;
  $distribution$ ::=
  | \code{Bernoulli} $\left(\;expr\;\code{,}\;expr\;\right)$
  | \code{Uniform}   $\left(\;expr\;\code{,}\;expr\;\right)$
  ;;
  $param$ ::=
  | $type\;amp\_opt\;identifier$
  ;;
  $amp\_opt$ ::=
  | \code{\&}
  | $\epsilon$
  ;;
  $param\_list$ ::=
  | $param\;param\_list\_tail$
  | $\epsilon$
  ;;
  $param\_list\_tail$ ::=
  | \code{,} $param\;param\_list\_tail$
  | $\epsilon$
\end{bnf}

\pagebreak

\begin{bnf}
  $expr$ ::=
  | $logical\_or\;prob\_assign\_opt$
  ;;
  $prob\_assign\_opt$ ::=
  | $prob\_assign$
  | $\epsilon$
  ;;
  $prob\_assign$ ::=
  | $prob\_choice\;expr$
  ;;
  $prob\_choice$ ::=
  | $\left\{\;expr\;\code{\:}\;expr\;\right\}$
  ;;
  $logical\_or$ ::=
  | $logical\_and\;logical\_or\_tail$
  ;;
  $logical\_or\_tail$ ::=
  | $\vert\vert\; logical\_and\;logical\_or\_tail$
  | $\epsilon$
  ;;
  $logical\_and$ ::=
  | $comparison\;logical\_and\_tail$
  ;;
  $logical\_and\_tail$ ::=
  | \code{\&\&} $comparison\;logical\_and\_tail$
  | $\epsilon$
  ;;
  $comparison$ ::=
  | $term\;comparison\_tail$
  ;;
  $comparison\_tail$ ::=
  | $comparison\_op\;term\;comparison\_tail$
  | $\epsilon$
  ;;
  $comparison\_op$ ::=
  | \code{>}
  | \code{>=}
  | \code{<}
  | \code{<=}
  | \code{==}
  | \code{!=}
  ;;
  $term$ ::=
  | $factor\;term\_tail$
  ;;
  $term\_tail$ ::=
  | \code{+} $factor\;term\_tail$
  | \code{-} $factor\;term\_tail$
  | $\epsilon$
  ;;
  $factor$ ::=
  | $division\;factor\_tail$
  ;;
  $factor\_tail$ ::=
  | \code{*} $division\;factor\_tail$
  | $\epsilon$
  ;;
  $division$ ::=
  | $unary\;division\_tail$
  ;;
  $division\_tail$ ::=
  | \code{/} $unary\;division\_tail$
  | \code{\%} $unary\;division\_tail$
  | $\epsilon$
  ;;
  $unary$ ::=
  | $unary\_op\;unary$
  | $primary$
  ;;
  $unary\_op$ ::=
  | \code{!}
  | \code{-}
  ;;
  $primary$ ::=
  | $bool$
  | $int\_literal$
  | $lval$
  | $\left(\;expr\;\right)$
  ;;
  $identifier$ ::=
  | $letter\;ident\_tail$
  ;;
  $ident\_tail$ ::=
  | $letter\_or\_digit\;ident\_tail$
  | \code{.} $ident\_tail$
  | $\colon\; ident\_tail$
  | \code{~} $ident\_tail$
  | $\epsilon$
  ;;
  $letter$ ::=
  | \code{a}
  | \code{b}
  | …
  | \code{Z}
  | \code{\_}
  ;;
  $digit$ ::=
  | \code{0}
  | …
  | \code{9}
  ;;
  $letter\_or\_digit$ ::=
  | $letter$
  | $digit$
  ;;
  $type$ ::=
  | \code{bool}
  | $int\_spec$
  | $int\_spec\;\left[\;int\_literal\;\right]$
  ;;
  $digit\_seq$ ::=
  | $digit\;digit\_seq$
  | $digit$
  ;;
  $int\_spec$ ::=
  | \code{u} $digit\_seq$
  | \code{s} $digit\_seq$
  ;;
  $int\_literal$ ::=
  | $digit\_seq\;int\_spec$
  ;;
  $bool$ ::=
  | \code{true}
  | \code{false}
\end{bnf}
